---
title: "Notes on air quality offset calculations"
author: "Christiaan Pauw, Nova Instituut"
date: "3 December 2015"
output: pdf_document
---
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(raster)
library(rasterVis)
library(knitr)
library(rgdal)

wd <- "/Users/christiaanpauw/Documents/Rpakette/AQoffset/"
setwd("/Users/christiaanpauw/Documents/Rpakette/AQoffset/")
rdir <-  file.path(paste(wd, "R/", sep=""))
datadir <- file.path(paste(wd, "../../AQ_DATA/Data/", sep = ""))

source(paste(rdir,'/count_exceed.R', sep=""))
source(paste(rdir, '/raster_dist_sum.R', sep=""))
source(paste(rdir, '/API.R', sep=""))
source(paste(rdir, 'knipNA.R', sep =""))
source(paste(rdir, 'raster_dist_plot.R', sep =""))
source(paste(rdir, 'pop2sp.R', sep =""))
source(paste(rdir, 'rasteriseCensus.R', sep =""))
source(paste(rdir, 'pointifyCensus.R', sep =""))
source(paste(rdir, 'project_boundary.R', sep =""))

new.proj <- "+proj=utm + south + zone=35 ellps=WGS84" 

load(paste(datadir, "kwaza_household_census.Rda", sep=""))
kwaN <- as.character(KWA@data$SP_NAME)
kwaCoor <- coordinates(KWA)
addkwa <- function(){layer(sp::sp.polygons(KWA, lwd = 0.25))}
txtkwa <- function(){layer(panel.text(x = coordinates(KWA)[,1], y = coordinates(KWA)[,2], label = as.character(KWA@data$SP_NAME), cex = 0.5))}
```

# Method applied to Kwazamokuhle with availible data 
```{r echo=FALSE, message=FALSE}
#remove the other datasets here to save memory
load(paste(datadir, "kwaza_eskom_24h_camx.Rda", sep = ""))
load(paste(datadir, "kwaza_eskom_year_camx.Rda", sep = ""))
load(paste(datadir, "kwaza_household_energyRaster.Rda", sep=""))
load(paste(datadir, "kwaza_API_hh.Rda", sep=""))
load(paste(datadir, "hh_24.Rda", sep=""))
load(paste(datadir, "EOP-fuel_users.Rda", sep=""))
kpext <- extent(KWA)
```

## Project boundary
The spatial extent of the project boundary is the extent of the ambient contribution of the managed facility above 2 ug/m3 per year or 19 ug/m3 per day in PM10 or SO2. 


### Extent of the impact of household emissions

#### Household emission sources

The results for main energy carrier for heating from the 2011 Census are give in the table and figure below.

```{r echo = FALSE}
kwa_census_df <- KWA@data[, c("SP_Name", "Electricity", "Gas", "Paraffin", "Wood", "Coal", "Animal.dung", "None")]
kwa_census_fuel_df <- data.frame(SP_NAME = KWA@data$SP_NAME, 
                                 Fuel = rowSums(kwa_census_df[,5:7], na.rm = T),
                                 No_fuel = rowSums(kwa_census_df[,c(2:4,8)], na.rm = T))
kable(kwa_census_df, row.names = FALSE, caption = "Main energy carrier for heating from the 2011 Census")
cat("\n")
```


```{r echo=FALSE}
levelplot(households[[-c(7:8)]], pretty = TRUE, par.settings=BuRdTheme, main = "Number of households by energy carrier \nused for heating: Census 2011", scales=list(draw=FALSE)) + addkwa()
```

It is clear that there are far fewer users of wood and dung than of coal. Because there are very few households who use wood that do not on occasion use coal as well, the responses for coal on the question "Mark ALL the energy carriers that you use for heating?" are used as an estimate for domestic solid fuel use. 
 
```{r echo = FALSE}
kable(kwa_census_fuel_df, row.names = FALSE, caption = "Solid fuel use for heating from the 2011 Census")
```

It is known however that because the Census asks only a question about the *main* energy carrier, the number of solid fuel users are underestimated. The results of the survey leads to a substantially higher estimate compared to the census. This is shown in the table below. The approximate spatial distribution of the households are also shown.  

```{r echo=FALSE}
kable(data.frame(coal.heat[!is.na(coal.heat[,2]), ], row.names = NULL))
```

```{r echo = FALSE}
pop2sp(coal.heat, raster = TRUE, refres = c(10,10), plot = TRUE, par.settings=BuRdTheme, main = "Number of coal users for heating: \nPoint estimate with 95% CI ", sub="'Yes' is coal use, 'No' is no coal use") + addkwa()
```

#### Application of thresholds to determine project boundary
The extent of the impact of household emissions in the baseline scenario above an annual average of 2ug/m3 or an daily average of 19ug/m3 is given below. Blank cells are outside of the project boundary

```{r echo = FALSE, message=FALSE}
raster_dist_plot(hh_24, mn = "Summary of modelled PM10 concentrations from household \ncoal burning in Kwazamokuhle", sb = "Daily averages over one year") + addkwa()
hh_24 <- project_boundary(r = hh_24) 
cat("\n")
raster_dist_plot(hh_24, mn = "Definition of the project boundary\nAreas where daily or annual PM10 exceed the threshold") + addkwa()
```

It is clear from the application of the threshold that if the modelling is correct, the impact of domestic burning is localised in close proximity to the emissions. The project boundary includes the whole of the main place Kwazamokuhle. Of all the sub-places that make up the main place Kwazamokuhle, the highest mean and maximum concentrations are found in Kwazamokuhle SP. 



## Baseline scenario
### Baseline emissions
Baseline emission are calculated from the results of a domestic fuel use survey. 

### Baseline states

The modeled baseline PM10 and SO2 resulting from household emissions in Kwazamokuhle is shown below. 
```{r echo = FALSE}
raster_dist_plot(hh_24, multi = c("pm10", "so2"), th=BuRdTheme) + addkwa()
```

The modeled baseline PM2.5, PM10 and SO2 from the industrial point source is shown below.

```{r echo=FALSE}
raster_dist_plot(crop(kwaza_eskom_24h_simple, kpext), multi = c("pm10","pm2.5", "so2")) + addkwa()
```

The count of days where the PM10 concentrations that resulted from household emissions and the industrial point source are modeled to exceed a specified level and the is shown below for the baseline scenario. 

```{r echo = FALSE}
hh_pm10_exceed = count_exceed(hh_24, pol = "pm10", min = 25, max = 200, by = 25, knip = TRUE)
levelplot(hh_pm10_exceed, main = "Count of days when PM10 from househols \nexceeded specified concentarion", sub = "182 days", par.settings=BuRdTheme) + addkwa()
```

```{r echo=FALSE}
#xl = expression(paste(mu,plain(g/m)^3))
bar_exceed(hh_pm10_exceed, ttl = "Aggregated of days when PM10 from households \nexceeded specified concentarion")
```

```{r echo = FALSE}
eskom_pm10_exceed = count_exceed(kwaza_eskom_24h_PM_SO2, pol = "pm10", min = 25, max = 200, by = 25, knip = TRUE)
levelplot(crop(eskom_pm10_exceed, kpext), main = "Count of days when PM10 from Eskom \nexceeded specified concentration", sub = "365 days", scales=list(draw=FALSE), par.settings=BuRdTheme) + addkwa()
```


```{r echo=FALSE}
bar_exceed(eskom_pm10_exceed, ttl = "Aggregated of days when PM10 from Eskom \nexceeded specified concentarion")
```

The count of days where SO2 concentration that resulted from household emissions is modeled to exceed a specified level is shown below for the baseline scenario. 

```{r echo=FALSE}
hh_so2_exceed = count_exceed(hh_24, pol = "so2", min = 25, max = 200, by = 25, knip = TRUE)
levelplot(hh_so2_exceed, main = "Count of days when SO2 from households \nexceeded specified concentration", sub = "182 days", par = BuRdTheme, scales=list(draw=FALSE)) + addkwa()
```

```{r echo=FALSE}
bar_exceed(hh_so2_exceed, ttl = "Aggregated count of days when SO2 from \nhouseholds exceeded specified concentration")
```

The count of days where SO2 from the industrial point source exceeded specified level for the baseline scenario is shown not in the same because all values are below the minimum level in the previous plot. 

### Baseline impact

The baseline impact represents the combined impact of all pollutants. A summary of the daily API resulting from households are given below.  

```{r echo=FALSE}
raster_dist_plot(kwaza_API, multi = NULL, mn = "Kwazamokuhle baseline API from households", sb = "182 days") + addkwa()
```

The baseline impact of the industrial point source for the full year is shown below.
```{r echo=FALSE}
raster_dist_plot(rasterAPI(crop(kwaza_eskom_24h_PM_SO2, kpext)), multi = NULL, mn = "Kwazamokuhle baseline API from households", sb = "182 days") + addkwa()
```

The count of days where the API that resulted from household emissions is modeled to exceed a specified level is shown below for the baseline scenario. 

```{r echo=FALSE}
levelplot(count_exceed(kwaza_API, min = 2, max = 10, by = 1, knip = TRUE) , par = BuRdTheme, scales=list(draw=FALSE), main = "Days in Kwazamokuhle where API \nfrom households exceeded specified level", sub = "182 days") + addkwa() 
```

## Project scenario

The project scanrio is the implementation on a stove exchage for full retrofit and LPG for all RDP houses who use coal. 

The estimates for fuel users per house type are shown below.

```{r echo = FALSE}
kable(fuel_house[!is.na(fuel_house[,2]), ])
fh = fuel_house[which(fuel_house[,2] == "Fuel_RDP"), ]
fh$place = unique(fuel_house[,1])[c(1,3:6)]
```

The target for the project activity is therefore between `r colSums(fh[,3:5])["Lower"]`  and `r colSums(fh[,3:5])["Upper"] ` with a point estimate of `r colSums(fh[,3:5])["PointEst"] `

```{r echo=FALSE}
kable(fh, caption = "Implementation targets per sub-place")
```

### Project emissions

#### Improvement per household

LPG 100%

Kitchen King = 0.018 * Union * 1.4    * 2 to be conservatve ~ 5%

#### Proportion of solid fuel using households reachable by the project



### Project states

### Project impacts





