---
title: "Air quality offset calculations for KwaZamokuhle"
author: "Christiaan Pauw, Nova Instituut"
date: "3 December 2015"
output: word_document

---

```{r global_options, echo=FALSE, message=FALSE, warning=FALSE}
library(raster)
library(rasterVis)
library(knitr)
library(rgdal)
library(pander)
knitr::opts_chunk$set(message=FALSE, warning=FALSE)


USER <- c("CJP", "AH", "BS")[2]

if (USER == "CJP") {
  wd <- "/Users/christiaanpauw/Documents/Rpakette/AQoffset/"
  setwd("/Users/christiaanpauw/Documents/Rpakette/AQoffset/")
  rdir <-  file.path(paste(wd, "R/", sep=""))
  datadir <- normalizePath(paste("Data/", sep = ""))
}

if (USER == "AH") {
  projdir <- "C:/Users/Alex H/Documents/R/AQoffset/"
  rdir <-  normalizePath(paste(projdir, "R/", sep=""))
  datadir <- normalizePath(paste(projdir, "Data/", sep = ""))
}

if (USER == "BS") {
  projdir <- "C:/Users/Willem/Desktop/AQOffset/AQoffset"
  rdir <-  normalizePath(paste(projdir, "R/", sep=""))
  datadir <- normalizePath(paste(projdir, "Data/", sep = ""))
}
  
source(paste(rdir, "/count_exceed.R", sep=""))
source(paste(rdir, '/raster_dist_sum.R', sep=""))
source(paste(rdir, '/API.R', sep=""))
source(paste(rdir, '/knipNA.R', sep =""))
source(paste(rdir, '/raster_dist_plot.R', sep =""))
source(paste(rdir, '/pop2sp.R', sep =""))
source(paste(rdir, '/rasteriseCensus.R', sep =""))
source(paste(rdir, '/pointifyCensus.R', sep =""))
source(paste(rdir, '/project_boundary.R', sep =""))
source(paste(rdir, '/SWI.R', sep =""))
source(paste(rdir, '/End.Point.Info.China.R', sep = ""))
source(paste(rdir, '/summarise.sicklist.R', sep = ""))
source(paste(rdir, '/rasterConcentrationResponse.R', sep = ""))

new.proj <- "+proj=utm + south + zone=35 ellps=WGS84" 

load(paste(datadir, "/kwaza_household_census.Rda", sep=""))
kwaN <- as.character(KWA@data$SP_NAME)
kwaCoor <- coordinates(KWA)
addkwa <- function() {latticeExtra::layer(sp::sp.polygons(KWA, lwd = 0.25))}
addkwa2 <- function() {layer(sp::sp.polygons(KWA, lwd = 10))}
txtkwa <- function() {layer(panel.text(x = coordinates(KWA)[,1], 
                                       y = coordinates(KWA)[,2], 
                                       label = as.character(KWA@data$SP_NAME), 
                                       cex = 0.5))}

```

```{r echo=FALSE, message=FALSE}

#remove the other datasets here to save memory
load(paste(datadir, "/kwaza_eskom_24h_camx.Rda", sep = ""))
load(paste(datadir, "/eskom.Rda", sep = ""))
#load(paste(datadir, "/kwaza_eskom_year_camx.Rda", sep = ""))
load(paste(datadir, "/kwaza_household_energyRaster.Rda", sep=""))
load(paste(datadir, "/kwaza_API_hh.Rda", sep=""))
load(paste(datadir, "/hh_24.Rda", sep=""))
load(paste(datadir, "/EOP-fuel_users.Rda", sep=""))
kpext <- extent(KWA)
eskom.base <- eskom * 0.15
```

```{r echo=FALSE}
# assumes that outcomesStack is a raster stack with 1 layer per outcome
sumIncidences <- function(outcomesStack) {
  dfIncidences <- data.frame(matrix(ncol = 2), 
                             stringsAsFactors = FALSE)
  names(dfIncidences) <- c("outcome", "total_cases")
  dfIncidences <- dfIncidences[-1,]
  
  for (i in 1:nlayers(outcomesStack)) {
    outcome <- tolower(names(outcomesStack)[i])
    totalIncidences <- sum(getValues(outcomesStack[[i]]), na.rm = TRUE)
    dfIncidences[nrow(dfIncidences) +1,] <- c(outcome, totalIncidences)
  }
  dfIncidences$outcome <- gsub(pattern = "cases_", 
                               replacement = "", 
                               x = dfIncidences$outcome, 
                               fixed = TRUE)
  dfIncidences$outcome <- gsub(pattern = "..", 
                               replacement = ".", 
                               x = dfIncidences$outcome, 
                               fixed = TRUE)
  dfIncidences$outcome <- gsub(pattern = ".", 
                               replacement = " ", 
                               x = dfIncidences$outcome, 
                               fixed = TRUE)
  
  dfIncidences$total_cases <- as.numeric(dfIncidences$total_cases)
  return(dfIncidences)
}
```

# Baseline scenario
## Baseline emissions
### Household emission sources

KwaZamokuhle is characterised by a high proportion of households who use coal for domestic cooking and heating. The importance of coal is visible in the results on the question on the main energy carrier for heating from the 2011 Census; these results are shown in the Table 1. 

```{r echo = FALSE, message=FALSE}
kwa_census_df <- KWA@data[, c("SP_Name", "Electricity", "Gas", "Paraffin", "Wood", "Coal", "Animal.dung", "None")]
kwa_census_fuel_df <- data.frame(SP_NAME = KWA@data$SP_NAME, 
                                 Fuel = rowSums(kwa_census_df[,5:7], na.rm = T),
                                 No_fuel = rowSums(kwa_census_df[,c(2:4,8)], na.rm = T))
kable(kwa_census_df, row.names = FALSE, caption = "Main energy carrier for heating from the 2011 Census")
cat("\n")
```

```{r echo=FALSE}
#levelplot(households[[-c(7:8)]], pretty = TRUE, par.settings=BuRdTheme, main = "Number of households by energy carrier \nused for heating: Census 2011", scales=list(draw=FALSE)) + addkwa()
```

It is clear that there are far fewer users of wood and dung than of coal. Because there are very few households who use wood that do not on occasion use coal as well, the responses for coal on the question "Mark ALL the energy carriers that you use for heating?" are used as an estimate for domestic solid fuel use. A summary is provided in Table 2. 
 
```{r echo = FALSE, message=FALSE}
kable(kwa_census_fuel_df, row.names = FALSE, caption = "Solid fuel use for heating from the 2011 Census")
```

It is known however that because the Census asks only a question about the *main* energy carrier, the number of solid fuel users are underestimated. The results of the survey lead to a substantially higher estimate compared to the census. The estimates of coal using households per sub-place derived from the household survey are shown in Table 3. It is clear that the number of coal using households derived from the survey results is substantially higher. 

```{r echo=FALSE, message=FALSE}
kable(data.frame(coal.heat[!is.na(coal.heat[,2]), ], row.names = NULL), caption = "Estimated number of coal using households per subplace with upper and lower bound of the 95% confidence interval")
```

The approximate spatial distribution of the households by coal use are shown below. The high concentration of coal users in the southern and eastern parts of KwaZamokuhle is clear.  

```{r , echo = FALSE, message=FALSE}
cat("\n")
pop2sp(coal.heat, 
       raster = TRUE, 
       refres = c(10,10), 
       plot = TRUE, 
       par.settings=BuRdTheme, 
       main = "Approximate location of coal users for heating: \nPoint estimate with 95% CI ", 
       sub="'Yes' is coal use, 'No' is no coal use") + addkwa()
cat("\n")
```

Baseline emissions are calculated from the results of a domestic fuel use survey. The estimates for fuel consumption based on the household survey are shown in Table 4. Winter coal consumption is understandably higher than summer consumption. Once again KwaZamokuhle SP has the highest consumption. 

```{r echo=FALSE, message=FALSE}
names(dta) <- gsub("suburb", "", names(dta))
names(dta) <- gsub("households", "#HH", names(dta))
names(dta) <- gsub("month.kg", "kg/m", names(dta))
names(dta) <- gsub("perc", "%", names(dta))
names(dta) <- gsub("winter.ave", "ave(W)", names(dta))
names(dta) <- gsub("summer.ave", "ave(S)", names(dta))
names(dta) <- gsub("\\.", " ", names(dta))
row.names(dta) <- NULL
pander(dta, split.cells = 40, 
       split.table = Inf, 
       rownames = NULL, 
       caption = "Baseline coal use by Suburb")
```

## Baseline states

The modelled baseline PM10 and SO~2~ resulting from household emissions in KwaZamokuhle is shown below. The baseline emissions from household coal use are summarised below. The tiles represent the 25th and 50th percentiles, the mean and the 75th and 99th percentiles. The last tile in each row represent the standard deviation. 

```{r echo = FALSE, message=FALSE}
raster_dist_plot(hh_24, multi = c("pm10", "so2"), th=BuRdTheme) + addkwa()
```

The distributions of the PM10 and SO~2~ are fairly similar but the concentration of PM10 is modelled to be higher. It is also clear that the concentration of both PM10 and SO~2~ decreases rapidly with distance from the household sources. 

\pagebreak
The modelled baseline secondary PM2.5 resulting from the Hendrina power station is shown below. The black dot indicates the position Kwazamokuhle.

```{r echo=FALSE}
raster_dist_plot(eskom.base, mn = "Secondary PM2.5 from Hendrina power station", sb = "364 days") + addkwa2()
```

The count of days where the PM10 concentrations that resulted from household emissions and the industrial point source is modelled to exceed a specified level, is shown below for the baseline scenario. 

```{r echo = FALSE}
hh_pm10_exceed = count_exceed(crop(hh_24, kpext), pol = "pm10", min = 75, max = 300, by = 50, knip = FALSE)
levelplot(hh_pm10_exceed, main = "Count of days when PM10 from househols \nexceeded specified concentarion", sub = "182 days", par.settings=BuRdTheme) + addkwa()
```

The exceedances of the daily PM10 standard related to household emissions occur over the southern and eastern side of KwaZamokuhle with the highest exceedance count at every level occurring in KwaZamokuhle SP.

```{r echo=FALSE}
#xl = expression(paste(mu,plain(g/m)^3))
# bar_exceed(hh_pm10_exceed, ttl = "Aggregated number of days when PM10 from households \nexceeded specified concentration", plot = FALSE)
concLevels <- names(hh_pm10_exceed)
df <- data.frame((cellStats(hh_pm10_exceed, sum) / 
                    (nrow(hh_pm10_exceed) * 
                     ncol(hh_pm10_exceed) * 
                     nlayers(hh_pm10_exceed))) *100 , stringsAsFactors = FALSE)
df$concentration <- concLevels
df[[3]] <- df[[1]]
df <- df[,2:3]
names(df) <- c("concentration", "%_days_in_exceedance")
rownames(df) <- NULL
kable(df)
```

```{r echo = FALSE}
# At this piont you can use setValues for populate the eskom raster with different data

eskom_kwa <- crop(eskom[[1:182]], extent(hh_24))
eskom_kwa <- disaggregate(eskom_kwa, c(nrow(hh_24),ncol(hh_24)))

eskom_kwa_base <- crop(eskom.base[[1:182]], extent(hh_24))
eskom_kwa_base <- disaggregate(eskom_kwa_base, c(36,41))

eskom_pm10_exceed = count_exceed(crop(eskom_kwa_base, kpext), pol = NULL, min = 0, max = 10, by = 0.1, knip = TRUE)
levelplot(eskom_pm10_exceed, main = "Count of days when PM2.5 from Eskom \nexceeded specified concentration", sub = "365 days", scales=list(draw=FALSE), par.settings=BuRdTheme) + addkwa()
```

```{r echo=FALSE}
#bar_exceed(eskom_pm10_exceed, ttl = "Aggregated of days when PM10 from Eskom \nexceeded specified concentration")
concLevels <- names(eskom_pm10_exceed)
df <- data.frame(cellStats(eskom_pm10_exceed, sum) / (nrow(eskom_pm10_exceed) * ncol(eskom_pm10_exceed) * nlayers(eskom_pm10_exceed)), stringsAsFactors = FALSE)
df$concentration <- concLevels
df[[3]] <- df[[1]]
df <- df[,2:3]
names(df) <- c("concentration", "%_days_in_exceedance")
rownames(df) <- NULL
kable(df)
```

\pagebreak
The count of days where SO~2~ concentration that resulted from household emissions is modelled to exceed a specified level, is shown below for the baseline scenario. 

```{r echo=FALSE}
hh_so2_exceed = count_exceed(crop(hh_24, kpext), pol = "so2", min = 100, max = 300, by = 25, knip = FALSE)
levelplot(hh_so2_exceed, main = "Count of days when SO2 from households \nexceeded specified concentration", sub = "182 days", par = BuRdTheme, scales=list(draw=FALSE)) + addkwa() + txtkwa()
```

Occurrences of exceedances of the daily SO~2~ standard related to coal combustion from households are less common than that for PM10. The area over which such occurrences are smaller than that of PM10 for every level and the count of exceedances for each level is also less. Once again most exceedances occur in KwaZamokuhle SP. 

```{r echo=FALSE}
#bar_exceed(hh_so2_exceed, ttl = "Aggregated count of days when SO2 from \nhouseholds exceeded specified concentration")
concLevels <- names(hh_so2_exceed)
df <- data.frame(cellStats(hh_so2_exceed, sum) / (nrow(hh_so2_exceed) * ncol(hh_so2_exceed) * nlayers(hh_so2_exceed)), stringsAsFactors = FALSE)
df$concentration <- concLevels
df[[3]] <- df[[1]]
df <- df[,2:3]
names(df) <- c("concentration", "%_observations_in_exceedance")
rownames(df) <- NULL
kable(df)
```

The count of days where SO~2~ from the industrial point source exceeded the specified level for the baseline scenario is not shown in the previous plot because all values are below the minimum level. 

### Combined baseline states

```{r echo=FALSE, message=FALSE}
hh_24[is.na(hh_24)] <- 0
base_PM10_vals <- getValues(hh_24[[183:364]]) + getValues(eskom_kwa_base)
allPM10_base <- hh_24[[183:364]]
allPM10_base <- setValues(allPM10_base, base_PM10_vals)
allSO2_base <- hh_24[[1:182]]
all_base <- stack(allPM10_base, allSO2_base)

```


# Project boundary
The spatial extent of the project boundary is the overlapping extent of the ambient contribution of the baseline emissions above 2 ug/m3 per year or 19 ug/m3 per day in PM10 or SO~2~, and the same for project emissions from the managed activity. Project emissions from the managed activity are the emissions from Eskom in the business-as-usual scenario. Baseline emissions from households are the emissions from households before implementation of the intervention. 

## Application of thresholds to determine project boundary

The baseline emissions from household coal use are summarised below. The tiles represent the 1st, 25th and 50th percentiles, the mean and the 75th and 99th percentiles. The last three tiles represent the standard deviation, inter-quartile range and 99% range of the data. 

```{r echo=FALSE, message=FALSE}
raster_dist_plot(all_base, multi = c("pm10", "so2"), mn = "Definition of the project boundary", sb = "Non-blank areas are where daily or \n annual PM10 exceed the threshold (182 days)", th=BuRdTheme) + addkwa() + txtkwa()
```

\pagebreak
The extent of the impact of household emissions in the baseline scenario above an annual average of 2ug/m3 or an daily average of 19ug/m3 is shown below. Blank cells are outside of the project boundary.

```{r echo = FALSE, message=FALSE}
masker <- project_boundary(r = hh_24, target_period = 'acute', only.mask = TRUE)
# Nie eskom nie want dan is daar nie 'n projke nie
#eskom_kwa_base <- project_boundary(r = eskom_kwa_base, target_period = 'acute', return.mask = FALSE)
all_base <- project_boundary(r = all_base, target_period = 'acute', return.mask = FALSE)

cat("\n")
raster_dist_plot(all_base, 
                 mn = "Definition of the project boundary", 
                 sb = "Non-blank areas are where daily or \n annual PM10 exceed the threshold (182 days)") + addkwa() + txtkwa()
```

It is clear from the application of the threshold that, if the modelling is correct, the impact of domestic burning is localised in close proximity to the emissions. The project boundary includes the whole of the main place KwaZamokuhle. Of all the sub-places that make up the main place KwaZamokuhle, the highest mean and maximum concentrations are found in KwaZamokuhle SP. 

## Baseline impact

Baseline impact can be determined using different calculation approaches. Four approaches will be demonstrated. These are: 
1.  Health risk approach
2.  Particle equivalence approach
3.  Standards weighted intake
4.  Burden of disease approach

### Health risk approach

The health risk approach uses an air quality index based on the relative risk of short term mortality associated with every pollutant. Here the pollutants are PM10 and SO~2~, but O~3~ and NO~2~ can potentially be added. The baseline impact represents the combined impact of all pollutants. A summary of the daily API resulting from households is given below.  

```{r echo=FALSE}
kwaza_API <- rasterAPI(all_base)
raster_dist_plot(kwaza_API, 
                 multi = NULL, 
                 mn = "Distribution of baseline API \nfor the baseline scenario", 
                 sb = "182 days") + addkwa() + txtkwa()
```

\pagebreak
The count of days where the baseline API is modelled to exceed a specified level is shown below for the baseline scenario. 

```{r echo=FALSE}
levelplot(count_exceed(kwaza_API, min = 3, max = 10, by = 1, knip = FALSE) , 
          par = BuRdTheme, 
          scales=list(draw=FALSE), 
          main = "Days in KwaZamokuhle where baseline API exceeded specified level", 
          sub = "182 days") + addkwa() + txtkwa()
```

KwaZamokuhle SP has the highest health risk related to short term exposure.  

### Particle equivalence approach

With sources that have a very localised dispersion, the particle equivalent impact (as PM10 equivalent) of that source is simply the concentration of the PM10 emitted by that source at every receptor; therefore, in the case of household emissions, the baseline impact is the same as the baseline state of PM10 shown above. 

### Standards weighted intake

The standards weighted intake approach is roughly equivalent to a population weighted air pollution index with the index weights determined by the standard. In this way it is related to the health risk approach that also requires population data. 

```{r echo=FALSE, message=FALSE}
swi_base <- SWI(all_base, pop = people)
levelplot(swi_base, par.settings = BuRdTheme, main = "Standards weighted intake in the baseline scenario", margin = FALSE) + addkwa() + txtkwa()
```


### Burden of disease approach

The burden of disease approach quantifies the actual or expected incidence of adverse health outcomes attributible to the exposure to ambient air pollution, and expresses the impact of the air pollution in terms of the proportion or number of cases of a specific outcome or as a weighted aggregate of such outcomes.

\pagebreak
The baseline burden of disease impact for PM10 resulting from household emissions is shown below.
```{r echo = FALSE, message=FALSE}
#hh_24
# create a people raster of the same extent as hh_24
households <- rasteriseCensus(KWA, ref = hh_24, refres = c(nrow(hh_24), ncol(hh_24)))
total_households <- calc(households, sum, na.rm = TRUE)
names(total_households) <- "Total_households"
total_households[total_households == 0] <- NA
people <- total_households*3.7
people[people==0] <- NA
names(people) <- "total"

# estimate incidences
rBoDPM10_hhE <- rasterCREP(sl = endlist,
                       pollutant = "PM10", 
                       cconc = hh_24[[grep(pattern = "pm10", 
                                           x = tolower(names(hh_24)), 
                                           fixed = TRUE)]], 
                       ppopr = people, 
                       bbase.conc = 10, 
                       rrisk.only = FALSE, 
                       verbose = FALSE, 
                       output = "AM",
                       ddelta = 1)

# summarise the number of cases across all days
rBoDPM10_hhE <- lapply(X = rBoDPM10_hhE, FUN = function(bOutcome) {
  return(calc(x = bOutcome, fun = sum))
})
rBoDPM10_hhE <- stack(rBoDPM10_hhE)

# tabulate total incidences per health outcome
kable(sumIncidences(rBoDPM10_hhE), digits = 2)

names(rBoDPM10_hhE) <- gsub("among.", "", names(rBoDPM10_hhE))
names(rBoDPM10_hhE) <- gsub("espiratory", "esp.", names(rBoDPM10_hhE))
names(rBoDPM10_hhE) <- gsub("hospital.admissions", "hosp.admin", names(rBoDPM10_hhE))
names(rBoDPM10_hhE) <- gsub("Cases_", "", names(rBoDPM10_hhE))
names(rBoDPM10_hhE) <- gsub("due.to.", "", names(rBoDPM10_hhE))
names(rBoDPM10_hhE) <- gsub("Cardiovascular", "CV", names(rBoDPM10_hhE))
names(rBoDPM10_hhE) <- gsub("all.ages", "all", names(rBoDPM10_hhE))

# plot incidences
levelplot(crop(rBoDPM10_hhE, kpext),
          main = "Burden of disease for PM10 
resulting from hh emissions,
given as number of cases per health outcome", 
          sub = "182 days",
          par = BuRdTheme) + addkwa() + txtkwa()
```

\pagebreak
The baseline burden of disease impact for SO~2~ resulting from household emissions is shown below.
```{r echo=FALSE, message=FALSE}
rBoDSO2_hhE <- rasterCREP(sl = endlist,
                      pollutant = "SO2", 
                      cconc = hh_24[[grep(pattern = "so2", 
                                           x = tolower(names(hh_24)), 
                                           fixed = TRUE)]], 
                      ppopr = people, 
                      bbase.conc = 10, 
                      rrisk.only = FALSE,
                      output = "AM",
                      verbose = FALSE,
                      ddelta = 1)

# summarise the number of cases across all days
rBoDSO2_hhE <- lapply(X = rBoDSO2_hhE, FUN = function(bOutcome) {
  return(calc(x = bOutcome, fun = sum))
})
rBoDSO2_hhE <- stack(rBoDSO2_hhE)

# tabulate total incidences per health outcome
kable(sumIncidences(rBoDSO2_hhE), digits = 2)

names(rBoDSO2_hhE) <- gsub("among.", "", names(rBoDSO2_hhE))
names(rBoDSO2_hhE) <- gsub("espiratory", "esp.", names(rBoDSO2_hhE))
names(rBoDSO2_hhE) <- gsub("hospital.admissions", "hosp.admin", names(rBoDSO2_hhE))
names(rBoDSO2_hhE) <- gsub("Cases_", "", names(rBoDSO2_hhE))
names(rBoDSO2_hhE) <- gsub("due.to.", "", names(rBoDSO2_hhE))
names(rBoDSO2_hhE) <- gsub("Cardiovascular", "CV", names(rBoDSO2_hhE))
names(rBoDSO2_hhE) <- gsub("all.ages", "all", names(rBoDSO2_hhE))

levelplot(crop(rBoDSO2_hhE, kpext),
          main = "Burden of disease for SO2 
resulting from hh emissions,
given as number of cases per health outcome",
          sub = "182 days",
          par = BuRdTheme) + addkwa() + txtkwa()
```

\pagebreak
The baseline burden of disease impact for PM10 resulting from the industrial point source is shown below.

```{r echo = FALSE, message=FALSE}
#kwaza_eskom_24h_simple
# create a people raster of the same extent as kwaza_eskom_24h_simple
households <- rasteriseCensus(KWA, 
                              ref = kwaza_eskom_24h_simple, 
                              refres = c(nrow(kwaza_eskom_24h_simple), 
                                         ncol(kwaza_eskom_24h_simple)))
total_households <- calc(households, sum, na.rm = TRUE)
names(total_households) <- "Total_households"
total_households[total_households == 0] <- NA
people <- total_households*3.7
people[people==0] <- NA
names(people) <- "total"
people <- crop(people, extent(all_base))

rBoDPM10 <- rasterCREP(sl = endlist,
                       pollutant = "PM10", 
                       cconc = all_base[[grep("pm10", names(all_base))]], 
                       ppopr = people, 
                       bbase.conc = 10, 
                       rrisk.only = FALSE, 
                       verbose = FALSE, 
                       output = "AM",
                       ddelta = 1)

# summarise the number of cases across all days
rBoDPM10_Ee <- lapply(X = rBoDPM10_Ee, FUN = function(bOutcome) {
  return(calc(x = bOutcome, fun = sum))
})
rBoDPM10_Ee <- stack(rBoDPM10_Ee)

# tabulate total incidences per health outcome
kable(sumIncidences(rBoDPM10_Ee), digits = 2)

levelplot(crop(rBoDPM10, kpext),
          main = "Burden of disease for PM10 
resulting in the baseline scenario emissions,
given as number of cases per health outcome",
          sub = "182 days",
          par = BuRdTheme) + addkwa() + txtkwa()
```

\pagebreak
The baseline burden of disease impact for SO2 resulting from the industrial point source is shown below.
```{r echo = FALSE, message=FALSE}
rBoDSO2_Ee <- rasterCREP(sl = endlist,
                       pollutant = "SO2", 
                       cconc = all_base[[grep("so2", names(all_base))]], 
                       ppopr = people, 
                       bbase.conc = 10, 
                       rrisk.only = FALSE, 
                       verbose = FALSE, 
                       output = "d",
                       ddelta = 1)

# summarise the number of cases across all days
rBoDSO2_Ee <- lapply(X = rBoDSO2_Ee, FUN = function(bOutcome) {
  return(calc(x = bOutcome, fun = sum))
})
rBoDSO2_Ee <- stack(rBoDSO2_Ee)

# tabulate total incidences per health outcome
kable(sumIncidences(rBoDSO2_Ee), digits = 2)

levelplot(crop(rBoDSO2, kpext),
          main = "Burden of disease for SO2 in the baseline scenario,
given as number of cases per health outcome",
sub = "182 days",
par = BuRdTheme) + addkwa() + txtkwa()

# tabulate total incidences per health outcome
kable(sumIncidences(rBoDSO2), digits = 2)
```

# Project scenario

The project scenario is the implementation of a stove exchange for all RDP houses who use coal, where the households exchange their old coal stoves for a full retrofit and LPG.

The estimates of fuel users per house type are shown below.

```{r echo = FALSE}
kable(fuel_house[!is.na(fuel_house[,2]), ])
fh = fuel_house[which(fuel_house[,2] == "Fuel_RDP"), ]
fh$place = unique(fuel_house[,1])[c(1,3:6)]
```

The target for the project activity is therefore between `r colSums(fh[,3:5])["Lower"]`  and `r colSums(fh[,3:5])["Upper"] ` with a point estimate of `r colSums(fh[,3:5])["PointEst"] `

```{r echo=FALSE}
kable(fh, caption = "Implementation targets per sub-place")
```

## Project emissions

### Improvement per household

LPG 100%

Kitchen King = 0.018 * Union * 1.4 * 2 to be conservative ~ 5%

#### Proportion of solid fuel using households reachable by the project

### Project states

#### Project states from households


```{r echo = FALSE}
raster_dist_plot(hh_24 * 0.49, multi =
                   c("pm10", "so2"), th=BuRdTheme) + addkwa()
```

#### Project states (all sources)

```{r echo=FALSE, message=FALSE}
# maak wees PS deur PS_hh, PS_es
hh_24_PS <- hh_24*0.49
proj_PM10_vals <- getValues(hh_24_PS[[183:364]]) + getValues(eskom_kwa)
allPM10_proj <- hh_24_PS[[183:364]]
allPM10_proj <- setValues(allPM10_proj, proj_PM10_vals)
allSO2_proj <- hh_24_PS[[1:182]]
all_proj <- stack(allPM10_proj, allSO2_proj)

raster_dist_plot(all_proj, multi = c("pm10", "so2"), mn = "Project states (all sources)", sb = "182 days") + addkwa()
```

```{r echo=FALSE, message=FALSE}
all_proj <- mask(crop(all_proj, extent(masker)), masker)
```

### Project impacts

#### Project standards weighted intake


```{r echo=FALSE, message=FALSE}
swi_proj <- SWI(all_proj, pop = people)
levelplot(swi_proj, par.settings = BuRdTheme, main = "Standards weighted intake in the project scenario", margin = FALSE) + addkwa() + txtkwa()
```

#### Projecth health risk approach

Project impacts  
```{r echo=FALSE, message=FALSE}
kwaza_API_proj <- rasterAPI(all_proj)
raster_dist_plot(kwaza_API_proj, 
                 multi = NULL, 
                 mn = "Distribution of baseline API \nfor the project scenario", 
                 sb = "182 days") + addkwa() + txtkwa()
```

#### Project burden of disease

```{r echo = FALSE, message=FALSE}
rBoDPM10_proj <- rasterCREP(sl = endlist,
                       pollutant = "PM10", 
                       cconc = all_proj[[grep("pm10", names(all_proj))]], 
                       ppopr = people, 
                       bbase.conc = 10, 
                       rrisk.only = FALSE, 
                       verbose = FALSE, 
                       out = "d")

levelplot(crop(rBoDPM10_proj, kpext),
          main = "Burden of disease for PM10 in the project scenario,
given as number of cases per health outcome",
sub = "182 days",
par = BuRdTheme) + addkwa() + txtkwa()

# tabulate total incidences per health outcome
kable(sumIncidences(rBoDPM10_proj), digits = 2)
```

## Impact reduction

#### Particle precursor approach

```{r echo=FALSE, message=FALSE}
PM_impact <- all_base[[grep("pm10", names(all_base))]] - all_proj[[grep("pm10", names(all_proj))]]
raster_dist_plot(PM_impact, mn = "Impact of the project activtiy on PM10 concentrations", sb = "182 days")

```

#### Health risk approach

```{r echo=FALSE, message=FALSE}
api_impact <- kwaza_API - kwaza_API_proj
raster_dist_plot(api_impact, 
                 multi = NULL, 
                 mn = "Distribution of API improvement", 
                 sb = "182 days") + addkwa() + txtkwa()
```

#### Standard weighted intake approach

```{r echo=FALSE, message=FALSE}
swi_impact <- swi_base - swi_proj
levelplot(swi_impact, main = "Impact in terms of reduction in standards weighted intake", sub = "Difference between baseline and project scenarion over 182 days", margin = FALSE, par.settings = BuRdTheme) + addkwa() + txtkwa()
```

#### Burden of disease approach 

```{r echo=FALSE, message=FALSE}
rBoDPM10_impact <- rBoDPM10 - rBoDPM10_proj
levelplot(rBoDPM10_impact, main = "Impact in terms of disease burden due to the \n implementation of the project activity", sub = "number of cases")
```

